Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers

### 2018-09-27 reports/preprint.tex
update preprint

In this work, a single tail volume was fitted to each lipid for all four surface pressures that were measured.
This is based on the assumption, that at all four surface pressures, the lipids adopt the same liquid expanded phase (as discussed Section \ref{sec:gixd}) and therefore any variation in the structure with surface pressure would manifest only as a change in the tail thickness, via the tail tilt angle.
It is clear when comparing Tables \ref{tab:water} and \ref{tab:liptab} that some of the tail volumes are reduced in the current XRR measurements compared to those determined previously.
The reduction was found to be \SIrange{8}{12}{\percent} for DPPC, DMPC and DLPC when compared with literature sources at \SIrange{24}{30}{\celsius}, this is in good agreement with the maximum compression percentage of \SI{15}{\percent} noted by Small \emph{et al.}\cite{Small1984}. DMPG shows a small increase in the tail volume relative to the literature value quoted at lower temperature.
Notably our value is similar to that found for DMPC on DES, which has the same tail structure and suggests that our results are at least self-consistent.

@@

The major difference between the two heads is the fact PG head is negatively charged whereas the PC head is zwitterionic (Figure \ref{fig:heads}).
It has been shown previously that, due to the electrostatic interaction, the conformation of the PC head is folded in water\cite{Gilliams2016}.
A similar structure may occur for the PG head, with a weaker interaction between the partially postively-charged alcoholic hydrogen atoms and the negatively-charged phosphate group.
Therefore, the observed increase found for the PG volume in DES when compared with water may be due to the unfolding of the PG head, where the solvent is capable of providing a greater screening effect to the PG head than that present in water.
This may not be observed for the PC head due to the greater strength of the folding interaction arising from the formally-charged nature of the ammonium group.
It would be anticipated that this unfolding would result in an increase in the thickness of the lipid head layer.
Previously, DPPG has been reported to have a head layer thickness of \SI[separate-uncertainty=true]{10.3\pm0.4}{\angstrom} at \SI{22}{\milli\newton\per\meter} from NR measurements\cite{Clifton2012}, which it slightly less than the \input{../output/dmpg/head.txt}\si{\angstrom} determined in the current work, further suggesting that the unfolding of the PG head.
Previously, the presence of short range, electrostatic interactions has been noted to impact the solution structure adopted by zwitterionic surfactants in choline chloride:glycerol\cite{Sanchez-Fernandez2018}.

>>>>

Due to the poor wetting properties of the DES, it was not possible to collect surface isotherm measurements.
However, grazing incidence X-ray diffraction was collected for DMPC and DPPC at \SI{30}{\milli\newton\per\meter}, indicating that the lipids had adopted the liquid condensed phase (see Section \ref{sec:gixd}).
This is the same phase behaviour as is observed in water \cite{Phillips1968} and by considering that there will be little interaction between the lipid tails and the DES it is unlikely that a significant change in the phase behaviour \cite{Nagle1976}.
The means that at all four surface pressures, as is the case on water, the lipids would adopt the same liquid condensed phase and any variation in the structure with surface pressure would manifest only as a change in the tail thickness, via the tail tilt angle.
This allowed a single tail volume to be fitted for each lipid at all four surface pressures that were measured.
It is clear when comparing Tables \ref{tab:water} and \ref{tab:liptab} that some of the tail volumes are reduced in the current XRR measurements compared to those determined previously.
The reduction was found to be \SIrange{8}{12}{\percent} for DPPC, DMPC and DLPC when compared with literature sources at \SIrange{24}{30}{\celsius}, this is in good agreement with the maximum compression percentage of \SI{15}{\percent} noted by Small \emph{et al.}\cite{Small1984}. DMPG shows a small increase in the tail volume relative to the literature value quoted at lower temperature.
Notably our value is similar to that found for DMPC on DES, which has the same tail structure and suggests that our results are at least self-consistent.

@@

The major difference between the two heads is the fact PG head is negatively charged whereas the PC head is zwitterionic (Figure \ref{fig:heads}).
It has been shown previously that, due to the electrostatic interaction, the conformation of the PC head is folded in water\cite{Gilliams2016}.
A similar structure may occur for the PG head, with a weaker interaction between the partially postively-charged alcoholic hydrogen atoms and the negatively-charged phosphate group.
Therefore, the observed increase found for the PG volume in DES when compared with water may be due to the unfolding of the PG head, where the solvent is capable of providing a greater screening effect to the PG head than that present in water.
This may not be observed for the PC head due to the greater strength of the folding interaction arising from the formally-charged nature of the ammonium group.
It would be anticipated that this unfolding would result in an increase in the thickness of the lipid head layer.
Previously, DPPG has been reported to have a head layer thickness of \SI[separate-uncertainty=true]{10.3\pm0.4}{\angstrom} at \SI{22}{\milli\newton\per\meter} \cite{Clifton2012} and \SI[separate-uncertainty=true]{9.7\pm1.0}{\angstrom} at \SI{15}{\milli\newton\per\meter} \cite{Ciumac2017} from NR measurements, which it slightly less than the \input{../output/dmpg/head.txt}\si{\angstrom} determined in the current work, further suggesting that the unfolding of the PG head.
Previously, the presence of short range, electrostatic interactions has been noted to impact the solution structure adopted by zwitterionic surfactants in choline chloride:glycerol\cite{Sanchez-Fernandez2018}.
